

%%% Choose a language %%%

\newif\ifEN
\ENtrue   % uncomment this for english
%\ENfalse   % uncomment this for czech

%%% Configuration of the title page %%%

\def\ThesisTitleStyle{mff} % MFF style
%\def\ThesisTitleStyle{cuni} % uncomment for old-style with cuni.cz logo
%\def\ThesisTitleStyle{natur} % uncomment for nature faculty logo

\def\UKFaculty{Faculty of Mathematics and Physics}
%\def\UKFaculty{Faculty of Science}

\def\UKName{Charles University in Prague} % this is not used in the "mff" style

% Thesis type names, as used in several places in the title
\def\ThesisTypeTitle{\ifEN BACHELOR THESIS \else BAKALÁŘSKÁ PRÁCE \fi}
%\def\ThesisTypeTitle{\ifEN MASTER THESIS \else DIPLOMOVÁ PRÁCE \fi}
%\def\ThesisTypeTitle{\ifEN RIGOROUS THESIS \else RIGORÓZNÍ PRÁCE \fi}
%\def\ThesisTypeTitle{\ifEN DOCTORAL THESIS \else DISERTAČNÍ PRÁCE \fi}
\def\ThesisGenitive{\ifEN bachelor \else bakalářské \fi}
%\def\ThesisGenitive{\ifEN master \else diplomové \fi}
%\def\ThesisGenitive{\ifEN rigorous \else rigorózní \fi}
%\def\ThesisGenitive{\ifEN doctoral \else disertační \fi}
\def\ThesisAccusative{\ifEN bachelor \else bakalářskou \fi}
%\def\ThesisAccusative{\ifEN master \else diplomovou \fi}
%\def\ThesisAccusative{\ifEN rigorous \else rigorózní \fi}
%\def\ThesisAccusative{\ifEN doctoral \else disertační \fi}



%%% Fill in your details %%%

% (Note: \xxx is a "ToDo label" which makes the unfilled visible. Remove it.)
\def\ThesisTitle{TaleCraft — Framework for 2D point-and-click adventure games}
\def\ThesisAuthor{Alžbeta Kulichová}
\def\YearSubmitted{2025}

% department assigned to the thesis
\def\Department{Department of Distributed and Dependable Systems}
% Is it a department (katedra), or an institute (ústav)?
\def\DeptType{Department}

\def\Supervisor{Mgr. Pavel Ježek, Ph.D.}
\def\SupervisorsDepartment{Department of Distributed and Dependable Systems}

% Study programme and specialization
\def\StudyProgramme{Computer Science}
\def\StudyBranch{Computer Graphics, Vision and Game Development }

\def\Dedication{
I would like to thank my supervisor Mgr. Pavel Ježek, Ph.D. for guidance he gave me during the writing of this thesis. I am also grateful to my family and friends for supporting me throughout my studies. Lastly, special thanks goes to my loving boyfriend, who was there for me even during the most difficult moments and without whom I would not be the person I am today.}

\def\AbstractEN{This thesis implements a C\# framework for 2D point-and-click adventure games built in the Unity engine. While similar tools exist in the Unity Asset store, point-and-click adventure games often rely on custom-built solutions, and there is still value in creating a lightweight, adaptable framework that can serve as both a development aid and a learning resource. It allows game developers to streamline the development of fundamental gameplay systems typical of the genre. At the same time, it offers flexibility and customizable components to accommodate a wide range of artistic styles and narrative structures. The framework provides core mechanics commonly found in point-and-click adventures such as a walking system, dialogue system, inventory system, and more. To showcase how the framework works in practice, a simple prototype game was created using its features. While the framework is not fully complete and does not include every mechanic a finished game might need, it provides a solid foundation and a starting point for future improvements and additional functionality.}

\def\AbstractCS{Tato práce implementuje C\# framework pro 2D point-and-click adventury v Unity enginu. Zatímco podobné nástroje existují v obchodě Unity Asset, dobrodružné hry typu point-and-click často spoléhají na řešení vytvořená na míru a stále je užitečné vytvořit lehký, přizpůsobivý framework, který může sloužit jako vývojová pomůcka i jako zdroj výuky. Umožňuje vývojářům her zefektivnit vývoj základních herních systémů typických pro tento žánr. Zároveň nabízí flexibilitu a přizpůsobitelné komponenty, aby vyhovovaly široké škále uměleckých stylů a narativních struktur. Framewok poskytuje základní mechaniky, které se běžně vyskytují v dobrodružstvích typu point-and-click, jako je systém pohybu, systém dialogů, systém inventáře a další. Abychom předvedli, jak framework funguje v praxi, byl vytvořen jednoduchý prototyp hry využívající jeho funkce. Přestože framework není zcela kompletní a nezahrnuje všechny mechaniky, které by hotová hra mohla potřebovat, poskytuje pevný základ a výchozí bod pro budoucí vylepšení a další funkce.}

% 3 to 5 keywords (recommended), each enclosed in curly braces.
% Keywords are useful for indexing and searching for the theses by topic.
\def\Keywords{%
{Unity} {framework} {2D game} {point-and-click adventure game}
}

% If your abstracts are long and do not fit in the infopage, you can make the
% fonts a bit smaller by this setting. (Also, you should try to compress your abstract more.)
% Alternatively, consider increasing the size of the page by uncommenting the
% geometry modification in thesis.tex.
\def\InfoPageFont{}
%\def\InfoPageFont{\small}  %uncomment to decrease font size

\ifEN\relax\else
% If you are writing a czech thesis, you additionally need to fill in the
% english translation of the metadata here!
\def\ThesisTitleEN{\xxx{Thesis title in English}}
\def\DepartmentEN{Department of Distributed and Dependable Systems}
\def\DeptTypeEN{Department}
\def\SupervisorsDepartmentEN{Department of Distributed and Dependable Systems}
\def\StudyProgrammeEN{Computer Science}
\def\StudyBranchEN{Computer Graphics and Game Engineering}
\def\KeywordsEN{%
\xxx{{key} {words}}
}
\fi
